Reduced plasticity and opportunity for selection in terrestrial ectotherm populations under climate change

### 2023-05-22 ms.qmd
Code mostly together but I need to do some careful checks as a few things are not quite making sense

# sigma2_v = typical sampling error variance**NEW STEP**
sub_data <- lnCVRQ10_data_wild[complete.cases(lnCVRQ10_data_wild$c_accl_time),]
  sigma2_v <- sum(1 / sub_data$V) * (dim(sub_data)[1] - 1) /
              (sum(1 / sub_data$V)^2 - sum((1 / sub_data$V)^2))

# Now check that var fix from raw data and remove sampling variance. **NEW STEP**
  v_fix2 <- var(sub_data$effect_size) - sigma2_v  #0.58

@@

@@

devtools::install_github("hughjonesd/ggmagnify") #https://github.com/hughjonesd/ggmagnify

@@

## **Discussion**

>>>>

# sigma2_v = typical sampling error variance
sub_data <- lnCVRQ10_data_wild[complete.cases(lnCVRQ10_data_wild$c_accl_time),]
  sigma2_v <- sum(1 / sub_data$V) * (dim(sub_data)[1] - 1) /
              (sum(1 / sub_data$V)^2 - sum((1 / sub_data$V)^2))

# Now check that var fix from raw data and remove sampling variance. 
  v_fix2 <- var(sub_data$effect_size) - sigma2_v  #0.58

@@

Projected changes in the opportunity for selection were highly variable across the globe, with some regions showing a decrease in the opportunity for selection, while others showed an increase (@fig-future). Overall, there was a net decrease in the opportunity for selection across the globe ()

@@

#devtools::install_github("hughjonesd/ggmagnify") #https://github.com/hughjonesd/ggmagnify

@@

```{r}
#| label: future_clim
#| echo: false
#| warning: false

# Now we want to take the predictions from `global` and use this for future climate sceneios. We have the future temperatures expected for each location. We can then use the predicted Q10 values to calculate the expected change in variance. We can then plot these globally.

# First, lets get the current and future temperature data for 2081-2085. We'll use the mean. 
  future_temp <- read.csv("./output/climate_data/gfdl2081-2085_air_and_SST_timeseries.csv")

# Lets clean this data up. We only need the mean temperature here, but this needs to be converted to Celcius as right now it's in Kelvins
  future_temp  <- future_temp  %>% 
                  dplyr::select(Species, Latitude, Longitude, mean)  %>% 
                  mutate(Celc_future = mean - 273.15,
                          match = interaction(Species, Latitude, Longitude))

# Current climate data
  climate <- read.csv("./output/climate_data/ERA5_air_and_SST_timeseries.csv")

# Clean up and create a match column. We need to get the last 5 years current to average temp and then convert this to Celcius
  climate2_future <- climate %>% mutate(match = interaction(Species, Latitude, Longitude)) %>% dplyr::select(match, Species, Latitude, Longitude, grep("X2018|X2019|X2020|X2021|X2022", colnames(climate)))  %>% mutate(mean = apply(.[, grep("X2018|X2019|X2020|X2021|X2022", colnames(.))], 1, function(x) mean(x, na.rm = TRUE)), Celc_curr = mean - 273.15) %>% dplyr::select(-grep("X2018|X2019|X2020|X2021|X2022", colnames(.)))

# Now we can merge teh climate datasets together so that we have the species, lat and long, current and future climate. Then we can calculate the expected 'increase' in T into the future

  climate_data <- left_join(climate2_future, future_temp, by = "match")  %>% dplyr::select(match, Latitude.x, Longitude.x, Species.x, Celc_curr, Celc_future)  %>% mutate(T_change = Celc_future - Celc_curr)  %>% rename(Species = Species.x, Latitude = Latitude.x, Longitude = Longitude.x)

# Lets look at future and current. Most, I would expect, should be above the line indicating higher future temperatures.
  ggplot(data = climate_data, aes(x = Celc_curr, y = Celc_future)) + geom_point() + geom_smooth(method = "lm") + labs(x = "Current Temperature (C)", y = "Future Temperature (C)") + theme_classic() + geom_abline(intercept = 0, slope = 1)

# Lets have a global plot of temperature changes to see it sptially 
  map2_temp_proj <- ggplot() +
    geom_map(
      data = world_map, map = world_map,
      aes(long, lat, map_id = region),
      color = "black", fill = "lightgray") +
    expand_limits(x = world_map$long, y = world_map$lat) +
    coord_map("moll") +
    geom_point(
      data = climate_data[complete.cases(climate_data$T_change),],
      aes(Longitude, Latitude, color = T_change),
      alpha = 0.7, size = 5
    ) + labs(x = "Longitude", y = "Latitude", color = TeX("$\\Delta$ Temp. ($^\\circ$C)"), title = "Current (2018-2022) vs Future (2080-2085)") + scale_colour_gradient2(trans = "reverse") + theme_map(base_size = 22) + theme(plot.title = element_text(hjust = 0.5))

# Now lets merge all_pred with the climate projections
  # Simplify columns 
  climate_data_new <- climate_data  %>% dplyr::select(-Latitude, -Longitude, -Species)

  # Mereg with model predcitions. This is the predcited lnCVRq10 for 10C change AT the specific latitude and longitude. We must assume that there is no plasticty or adaptation at each site to make projections
  all_pred2 <- all_pred %>% mutate(match = interaction(spp, Latitude, Longitude)) %>% 
              left_join(., climate_data_new, by = "match")  %>% 
              mutate(lnCVR_future = (Estimate / 10) * T_change,
                      lnCVR_F_l95 = (Q2.5 / 10) * T_change,
                      lnCVR_F_u95 = (Q97.5 / 10) * T_change, 
                      per_change = (exp(lnCVR_future) - 1) * 100,
                      per_change_l95 = (exp(lnCVR_F_l95) - 1) * 100,
                      per_change_u95 = (exp(lnCVR_F_u95) - 1) * 100)

# Need to take a step back
  # 1. Why do we have NA in this data set? 
  # 2. Why are l95 and u95 for marine not working
  # 3. Apply to the posteriors, rather than summary.

tables_future <- all_pred2 %>% group_by(habitat) %>% summarise(mean = mean(per_change, na.rm = TRUE), 
                                                                l95 = mean(per_change_l95, na.rm = TRUE), 
                                                                u95 = mean(per_change_u95, na.rm = TRUE))

```{r}
#| label: fig-future
#| echo: false
#| warning: false

map2_future <- ggplot() +
    geom_map(
      data = world_map, map = world_map,
      aes(long, lat, map_id = region),
      color = "black", fill = "lightgray") +
    expand_limits(x = world_map$long, y = world_map$lat) +
    coord_map("moll") +
    geom_point(
      data = all_pred2[complete.cases(all_pred2$per_change),],
      aes(Longitude, Latitude, color = per_change),
      alpha = 0.7, size = 5
    ) + labs(x = "Longitude", y = "Latitude", color = TeX("% $\\Delta$"), title = "Future (2080-2085)") + scale_colour_viridis() + theme_map(base_size = 22) + theme(plot.title = element_text(hjust = 0.5))
```
## **Discussion**
